---
title: 'PSTAT 174 Final Project: Milk Production'
author: "Chaewon Shin"
date: Spring 2020
output:
  pdf_document: default
  html_document: default
---

\newpage

# Introduction

## Data Preparation
To prepare building the model, we first create a training data that will be used to build the model. The test data will be used in to forecast the data. 
```{r}
#Setting working directory
library(fma)
library(MASS)
library(astsa)
library(qpcR)
```
```{r}
#Leaving 12 data points for validation 
milkt <- milk[c(1:144)] #Training dataset U to build a model
milktest <- milk[c(145:156)] #Test dataset to forecast (last year)

```

## Plotting the time series (Raw Data)
Based on the raw data, some of the immediate observations that can be made is that the data is highly non-stationary, has a linear trend, and has seasonality. The non-constant variance and mean implies that stabilization is needed. In order to remove trend, we will perform differencing, and to stabilize variance, we will perform transformation.
```{r}
# Raw Data Time Series Plot
tsdat <- ts(milk, start = c(1962,1), end = c(1974,2), frequency = 12)
ts.plot(tsdat, main = "Raw Data: Milk production")
```

## Training data set

```{r}
#Plotting the training data
plot.ts(milkt, main = "Model Training Data: Milk Production")
#Adding the trend line to plot
fitt <- lm(milkt ~ as.numeric(1:length(milkt)))
abline(fitt, col = 'red')
#Adding mean to the plot
abline(h=mean(milk), col = 'blue')

```

## Confirming non-stationarity of data
Before performing transformations and differencing, we will confirm the nonstationarity of the data by plotting the histogram and the acf. From the histogram, we can see that the data is skewed slightly to the left, and from the ACF graph, we can see that the ACF remains large and periodic. 
```{r}
#Plotting the histogram
hist(milkt, col="light blue", xlab="", main="Histogram of milk data")
#Plotting the acf
acf(milkt,lag.max=40, main="ACF of the Milk Data")
```

## Transformations
To determine which transformation to perform, we will use a Box-Cox transformation to determine lambda. 
```{r}
#Box-Cox Transformation
bcTransform <- boxcox(milkt~ as.numeric(1:length(milkt))) 
lambda = bcTransform$x[which(bcTransform$y == max(bcTransform$y))]
milkt.bc = (1/lambda)*(milkt^lambda-1)
hist(milkt.bc, col="light blue", xlab="", main="histogram; bc(U_t)")
```
From the Box-Cox plot, we can see that the confidence interval includes lambda = 0, which implies the Box-Cox transformation is given by Yt = logXt. Here, the value of lambda is close enough to 0 to proceed with a log transformation.

```{r}
#Log Transformation
milkt.log = log(milkt)
plot.ts(milkt.log)
hist(milkt.log, col="light blue", xlab="", main="histogram; ln(U_t)")
```


## Decomposition of data
```{r}
#Installing necessary packages
library(ggplot2)
library(ggfortify)
y <- ts(as.ts(milkt.log), frequency = 12)
decomp <- decompose(y)
plot(decomp)
```
From the decomposition, we can see that there is a upward trend as well as a seasonal component. Our next step would be to remove seasonality, which is done by differencing. 

## Differencing
```{r}
#Initial variance of milkt
var(milkt)
#Differencing at lag 12
milkt.12 <- diff(milkt, lag=12)
plot.ts(milkt.12, main="U_t differenced at lag 12")
var(milkt.12)
abline(h=mean(milkt.12), col="blue")
fit <- lm(milkt.12 ~ as.numeric(1:length(milkt.12)))
abline(fit, col="red") 
```
By differencing at lag 12, the variance has gone down significantly. 
```{r}
#Differencing again at lag 1
milkt.data <- diff(milkt.12, lag= 1)
plot.ts(milkt.data, main = "U_t differenced at lag 12 and lag 1")
var(milkt.data)
abline(h=mean(milkt.data), col="blue")
fit.1 <- lm(milkt.data ~ as.numeric(1:length(milkt.data)))
abline(fit.1, col="red") 
```
By differencing again at lag 1, the varaince has gone down, and most of the seasonality has been removed. The data now appears to be stationary, but to check further, we will look at the ACF.

## ACF of Data
```{r}
acf(milkt, lag.max=40, main="ACF of the U_t")
acf(milkt.12, lag.max=40, main="ACF of the U_t, differenced at lag 12")
acf(milkt.data, lag.max=40, main="ACF of the U_t, differenced at lags 12 and 1")
hist(milkt.data, col="light blue", xlab="", main="Histogram of U_t differenced at lags 12 & 1")

```

## Plotting histogram with normal curve
```{r}
hist(milkt.data, col="light blue", xlab="", main="Histogram of U_t differenced at lags 12 & 1", prob=TRUE)
m<-mean(milkt.data)
std<- sqrt(var(milkt.data))
curve( dnorm(x,m,std), add=TRUE )

```
```{r}
hist(milkt, col="light blue", xlab="", main="Histogram of milk data")
mt<-mean(milkt)
stdt<- sqrt(var(milkt))
curve( dnorm(x,mt,stdt), add=TRUE )
```


## ACF and PACF Analysis
```{r}
acf(milkt.data, lag.max=50, main="ACF of the U_t, differenced at lags 12 and 1")
pacf(milkt.data, lag.max=50, main="PACF of the U_t, differenced at lags 12 and 1")
```
- Since we applied one seasonal differencing, D = 1 at lag s = 12
- The ACF shows a strong peak h = 1s, and smaller peaks at h = 2s, 3s. A possible choice for the MA part is Q = 1 or Q = 3. (Q=1)
- The PACF shows a strong peak at h = 1s, and smaller peaks at h = 2s, 3s. A possible choice for AR part is P = 1 or P = 4.
- To remove the trend, we applied one difference, so d = 1.
- The ACF appears to be tailing off, so a possible choice for the MA part is q = 0.
- The PACF cuts off at h = 1s, so a possible choice for AR is p = 0. 

We will be testing the following models: 
i. SARIMA (p=0, d=1, q=0) x (P=1, D=1, Q=1) s=12
ii. SARIMA (p=0, d=1, q=0) x (P=4, D=1, Q=1) s=12
iii. SARIMA (p=0, d=1, q=0) x (P=1, D=1, Q=3) s=12
iv. SARIMA (p=0, d=1, q=0) x (P=4, D=1, Q=3) s=12

## Testing the models
```{r}
# Fitting the candidate models (model i)
fit.i <- sarima(xdata = milk, p = 0, d = 1, q = 0, P = 1, D = 1, Q = 1, S = 12, details = F)
fit.i$fit$coef
plot(fit.i$fit$residuals)

# Fitting the candidate models (model ii)
fit.ii <- sarima(xdata = milk, p = 0, d = 1, q = 0, P = 4, D = 1, Q = 1, S = 12, details = F)
fit.ii$fit$coef
plot(fit.ii$fit$residuals)

# Fitting the candidate models (model iii)
fit.iii <- sarima(xdata = milk, p = 0, d = 1, q = 0, P = 1, D = 1, Q = 3, S = 12, details = F)
fit.iii$fit$coef
plot(fit.iii$fit$residuals)

# Fitting the candidate models (model iv)
fit.iv <- sarima(xdata = milk, p = 0, d = 1, q = 0, P = 4, D = 1, Q = 3, S = 12, details = F)
fit.iv$fit$coef
plot(fit.iv$fit$residuals)
```

```{r}
#Testing SMA Models (Q = 1 or 3, q = 0)

#Model i
arima(milkt, order=c(0,1,0), seasonal = list(order = c(0,1,1), period = 12), method="ML")
AICc(arima(milkt, order=c(0,1,0), seasonal = list(order = c(0,1,1), period = 12), method="ML"))

#Model iii
arima(milkt, order=c(0,1,0), seasonal = list(order = c(0,1,3), period = 12), method="ML")
AICc(arima(milkt, order=c(0,1,0), seasonal = list(order = c(0,1,3), period = 12), method="ML"))

#Model iii w/ zero coefficients
arima(milkt, order=c(0,1,0), seasonal = list(order = c(0,1,3), period = 12), fixed = c(NA, 0, 0), method="ML")
AICc(arima(milkt, order=c(0,1,0), seasonal = list(order = c(0,1,3), period = 12), fixed = c(NA, 0, 0), method="ML"))

arima(milkt, order=c(0,1,0), seasonal = list(order = c(0,1,3), period = 12), fixed = c(NA, NA, 0), method="ML")
AICc(arima(milkt, order=c(0,1,0), seasonal = list(order = c(0,1,3), period = 12), fixed = c(NA, NA, 0), method="ML"))

arima(milkt, order=c(0,1,0), seasonal = list(order = c(0,1,3), period = 12), fixed = c(NA, 0, NA), method="ML")
AICc(arima(milkt, order=c(0,1,0), seasonal = list(order = c(0,1,3), period = 12), fixed = c(NA, 0, NA), method="ML"))
```
Here, the SMA model with Q = 3, q = 0 has a lower AICc value. 

```{r}
#Including the AR component Models (P = 1 or 4, p = 0)

#Model i
arima(milkt, order=c(0,1,0), seasonal = list(order = c(4,1,1), period = 12), method="ML")
AICc(arima(milkt, order=c(0,1,0), seasonal = list(order = c(4,1,1), period = 12), method="ML"))

arima(milkt, order=c(0,1,0), seasonal = list(order = c(1,1,1), period = 12), method="ML")
AICc(arima(milkt, order=c(0,1,0), seasonal = list(order = c(1,1,1), period = 12), method="ML"))

arima(milkt, order=c(0,1,0), seasonal = list(order = c(4,1,1), period = 12), fixed = c(NA, NA, NA, NA, 0), method="ML")
AICc(arima(milkt, order=c(0,1,0), seasonal = list(order = c(4,1,1), period = 12), fixed = c(NA, NA, NA, NA, 0), method="ML"))

arima(milkt, order=c(0,1,0), seasonal = list(order = c(1,1,1), period = 12), fixed = c(0, NA), method="ML")
AICc(arima(milkt, order=c(0,1,0), seasonal = list(order = c(1,1,1), period = 12), fixed = c(0, NA),method="ML"))

```

## Final models to test
```{r}
#modelA <- arima(milkt, order=c(0,1,0), seasonal = list(order = c(4,1,1), period = 12), method="ML")
#modelA
AICc(arima(milkt, order=c(0,1,0), seasonal = list(order = c(4,1,1), period = 12), method="ML"))

#modelB <- arima(milkt, order=c(0,1,0), seasonal = list(order = c(4,1,1), period = 12), fixed = c(NA, NA, NA, NA, 0), method="ML")
#modelB
AICc(arima(milkt, order=c(0,1,0), seasonal = list(order = c(4,1,1), period = 12), fixed = c(NA, NA, NA, NA, 0), method="ML"))
```

```{r}
#Determining which coefficients to fix to 0
confint(modelA)
confint(modelB)
```

Here, by adding the AR component, the AICc value decreases when p = 1 and P = 3. 

# Model Diagnostics
```{r}

fit1 <- sarima(xdata = milkt, p = 0, d = 1, q = 0, P = 1 , D = 1, Q = 1, S = 12, details = F)
fit1
fit1f <- sarima(xdata = milkt, p = 0, d = 1, q = 0, P = 1 , D = 1, Q = 1, S = 12, fixed = c(0,NA), details = F)
fit1f

fit2 <- sarima(xdata = milkt, p = 0, d = 1, q = 0, P = 4 , D = 1, Q = 1, S = 12, details = F)
fit2
fit2f <- sarima(xdata = milkt, p = 0, d = 1, q = 0, P = 4 , D = 1, Q = 1, S = 12, fixed = c(NA, NA, NA, NA, 0), details = F)

fit3 <- sarima(xdata = milkt, p = 0, d = 1, q = 0, P = 1 , D = 1, Q = 3, S = 12, details = F)
fit3

fit4 <- sarima(xdata = milkt, p = 0, d = 1, q = 0, P = 4 , D = 1, Q = 3, S = 12, details = F)
fit4

```

```{r}
#FINAL MODELS:
modelA <- sarima(xdata = milkt, p = 0, d = 1, q = 0, P = 1 , D = 1, Q = 1, S = 12, details = F) #fit2f
modelA
modelB <- sarima(xdata = milkt, p = 0, d = 1, q = 0, P = 0 , D = 1, Q = 1, S = 12, details = F) #fit5
modelB 
```
The models are given by: 
  A. SARIMA (0,1,0)x(1,1,1)s=12
    \[(1-0.0801B^{12})Y_t = (1-0.7387B^{12})Z_t\]
  B. SARIMA (0,1,0)x(0,1,1)s=12
    \[Y_t = (1-0.6877B^{12})Z_t\]
## Checking Stationarity and Invertibility

For model A, we have an SAR model. Since all AR models are invertible, model A in invertible. To check stationarity, we will check the roots:
For model B, we have a SMA model. Since all MA models are stationary, model B is stationary. 
To check invertibility: Model B is invertible since |theta1| < 1

```{r}
#Polyroot function
polyroot(c(1,-.68771))
```



#Diagnostic Checking
```{r}
#MODEL A
fitA <- arima(milkt, order=c(0,1,0), seasonal = list(order = c(1,1,1), period = 12), method="ML")
resA <- residuals(fitA)
hist(resA, density = 15, breaks=20, col="dark blue", xlab="", main = "Histogram of Model A residuals", prob=TRUE)
mA <- mean(resA)
stdA <- sqrt(var(resA))
curve(dnorm(x,mA,stdA), add=TRUE )
plot.ts(resA, main = "Plot of Model A residuals")

#MODEL B
fitB <- arima(milkt, order=c(0,1,0), seasonal = list(order = c(0,1,1), period = 12), method="ML")
resB <- residuals(fitB)
hist(resB,density = 15,breaks=20, col="blue", xlab="", main = "Histogram of Model B residuals", prob=TRUE)
mB <- mean(resB)
stdB <- sqrt(var(resB))
curve(dnorm(x,mB,stdB), add=TRUE )
plot.ts(resB, main = "Plot of Model B residuals")
```

```{r}
#Normal Q-Q Plot
plot.ts(resA, main = "Plot of Model A residuals")
fittA <- lm(resA ~ as.numeric(1:length(resA)))
abline(fittA, col="red")
abline(h=mean(resA), col="blue")
qqnorm(resA,main= "Normal Q-Q Plot for Model A")
qqline(resA,col="blue")

plot.ts(resB, main = "Plot of Model B residuals")
fittB <- lm(resB ~ as.numeric(1:length(resB)))
abline(fittB, col="red")
abline(h=mean(resB), col="blue")
qqnorm(resB,main= "Normal Q-Q Plot for Model B")
qqline(resB,col="blue")
```


```{r}
#ACF/PACF of residuals
acf(resA, lag.max = 50)
pacf(resA,lag.max = 50)

acf(resB,lag.max = 50)
pacf(resB,lag.max = 50)
```

```{r}
#Tests
shapiro.test(resA)
Box.test(resA, lag = 12, type = c("Box-Pierce"), fitdf = 0)
Box.test(resA, lag = 12, type = c("Ljung-Box"), fitdf = 0)
Box.test(resA^2, lag = 12, type = c("Ljung-Box"), fitdf = 0)

shapiro.test(resB)
Box.test(resB, lag = 12, type = c("Box-Pierce"), fitdf = 0)
Box.test(resB, lag = 12, type = c("Ljung-Box"), fitdf = 0)
Box.test(resB^2, lag = 12, type = c("Ljung-Box"), fitdf = 0)
```

# Forecasting
```{r}
library(forecast)
fitA <- arima(milkt, order=c(0,1,0), seasonal = list(order = c(1,1,1), period = 12), method="ML")
forecast(fitA)


#Plotting the forecast
pred.tr <- predict(fitA, n.ahead = 12)
U.tr= pred.tr$pred + 2*pred.tr$se 
L.tr= pred.tr$pred - 2*pred.tr$se
ts.plot(milkt, xlim=c(1,length(milkt)+12), ylim = c(min(milkt),max(U.tr)), main = "Plot of forecasted values")
lines(U.tr, col="blue", lty="dashed")
lines(L.tr, col="blue", lty="dashed")
points((length(milkt)+1):(length(milkt)+12), pred.tr$pred, col="red")
```

```{r}
#Forecast on original data
ts.plot(milkt, xlim = c(100,length(milkt)+12), ylim = c(250,max(U.tr)), main = "Forecasted Values")
lines(U.tr, col="blue", lty="dashed")
lines(L.tr, col="blue", lty="dashed")
points((length(milkt)+1):(length(milkt)+12), pred.tr$pred, col="red")

ts.plot(milk[c(1:156)], xlim = c(100,length(milkt)+12), ylim = c(250,max(U.tr)), main = "Forecasted Values on Original Data")
lines(U.tr, col="blue", lty="dashed")
lines(L.tr, col="blue", lty="dashed")
points((length(milkt)+1):(length(milkt)+12), pred.tr$pred, col="red")

```

